---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Reproducible Research Assignment 1

## Synopsis

This assignment makes use of data from a personal activity monitoring device, such as Fitbit, Nike Fuelband or Jawbone up. The data is collected throughout the day, in 5 min inetrvals. For the assigment we will analyze 2 months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Data

*Datset:* [Activity Monitoring Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)  
*Variables:*  
    1. **steps:** Number of steps taking in a 5-minute interval (missing values are coded as NA)   
    2. **date:** The date on which the measurement was taken in YYYY-MM-DD format   
    3. **interval:** Identifier for the 5-minute interval in which measurement was taken  

### Loading Preprocessing the data into R
```{r, echo=TRUE}
# Set the working directory
setwd("D:/DataScience/Reproducible Research/Assignment 1")

# Read the activity data
activity_data <- read.csv('activity.csv')
head(activity_data)

#Remove NA from the activity file
clean_activity_data <- na.omit(activity_data)
head(clean_activity_data)
```

## Analysis of the data
### What is mean total number of steps taken per day?

Steps followed:  
 1. Calculate the total number of steps taken per day  
 2. Make a histogram of the total number of steps taken each day  
 3. Calculate and report the mean and median of the total number of steps taken per day
 
```{r, echo=TRUE}
# Aggregate steps in a day 
Agg_Step <- aggregate(steps ~ date, clean_activity_data, sum)

# Plot the histogram 
hist(Agg_Step$steps, main ="Total Number of steps taken per day", xlab = "Steps/Day", col = "blue")

#Get the mean total number of steps per day
mean(Agg_Step$steps)

#Get the median total number of steps per day
median(Agg_Step$steps)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
